---
title: "Analysis of 2018-2020 Eden Landing data using old MEM-PEPRMT"
author: "Michael Najarro"
date: "11/17/2020"
output: rmarkdown::github_document
---


```{r}
library(pacman)
p_load(oldrCTM331,
       tidyverse,
       magrittr,
       tinytex,
       R.matlab,
       here,
       dplyr,
       tidyr,
       tibble,
       tinytex,
       knitr,
       tictoc,
       beepr)
```


# *Phase 1: MEM*

### *Step 1: Run MEM with Cohorts*

```{r}
#import elevation data to environment.
elevation<- read.table(file="./Data_sets/EDENLAND082620.txt",sep=",")

a <- elevation %>%
  rename(record=V1,
         S1=V2,
         S2=V3,
         elevation=V4,
         edcv=V5) %>%
  slice(-25) %>%
  summarise(elev = mean(elevation)*100) %>%
  pull(.)
```


```{r}
omPackingDensity <- 0.82

memCohortExample2 <- runMemWithCohorts(startYear=2010,
                  rslrT1=0.09, #.9
                  rslrTotal=250,
                  initElv=a,
                  MSL=98.1,
                  MHW=194.7,
                  MHHW=213.7,
                  MHHWS= 287,
                  ssc= .000020,#.000025,#.000030,
                  lunarNodalAmp=2.5,
                  bMax=.90,#0.84,#.90,#.048,
                  zVegMin=100,
                  zVegMax=350,#250,#original350,
                  zVegPeak=180,                  plantElevationType="orthometric",
                  rootToShoot=2,
                  rootTurnover=0.5,
                  rootDepthMax=30,
                  shape="linear",
                  omDecayRate=0.8,
                  recalcitrantFrac=0.35,#0.2,
                  settlingVelocity=0.275,
                  omPackingDensity=0.82,
                 rootPackingDensity = omPackingDensity,
            mineralPackingDensity=2.43,
                  coreYear = 2050#,
                  #coreDepth = 50
                 # coreMins =coreMins[c(1:26),],
                 # coreMaxs = coreMaxs[c(1:26),]
)


slow <- tapply(memCohortExample2$cohorts$slow_OM,
           memCohortExample2$cohorts$year,
           FUN = mean)
slow
fast <- tapply(memCohortExample2$cohorts$fast_OM,
           memCohortExample2$cohorts$year,
           FUN = mean)
fast
resp <-tapply(memCohortExample2$cohorts$respired_OM,
           memCohortExample2$cohorts$year,
           FUN = mean)
resp
```



## **Step 2: obtain the sum of slow and fast organic matter pools**

At this point I entirely skip the calculation and downscaling of NPP.

Instead I calculate two important values from MEM to input into the respiration component of PEPRMT. These values will directly correspond to the slow and fast organic matter components of each soil cohort from MEM. The sum of these values approximate the carbon stored within the labile and SOC pools described in the PEPRMT model.


The MEM ouputs for carbon -The slow and fast organic measures - will become inputs for the ecosystem respiration component of PEPRMT. Specifically, they will equate to the sum of slow and fast organic matter that is within one meter below the soil surface at and up to the time of interest (includes all years prior, up to, and include year of interest) in accordinance with the PEPRMT model (see Oikawa 2017) to simulate soil sampling of core data. 

Note thatin previous iterations of MEM-PEPRMT in Rush Ranch, we did not consider the 1 meter depth because the total height of the soil cohorts from MEM were within 1 meter.

The relationship between these pools can be described as follows:

```{r}
y<- tribble(
  ~MEM_Variable, ~Equiv._PEPRMNT_C_Pool, ~PEPRMNT_Variable, ~theta_element,
  "Slow_OM", "Soil Organic Carbon (SOC) Pool", "S2", "theta[5]",
  "Fast_OM", "Labile Carbon Pool", "S1", "theta[6]"
)

kable(y)
```


Here I calculate the depth of the soil accreted for 2018 and filter out the 2018 data from MEM outputs. I also check if the soil is within one meter of the surface.

```{r}
# first identify the values of slow and fast om
# that are: 
#   1. within a meter
#   2. within a time interval of the earliest
#       year up to, and including the year of 
#       interest.

#note:
# Within a given year, all rows before the newest
# incremented "age" (n+1th row) are considered past cohorts,
# so only sum the rows within a year of interest!

#2.a) create table cohort up to year 2017.
cohort <- as.data.frame(memCohortExample2$cohorts)
#cores <-- as.data.frame(memCohortExample2$core)

#2.b) create a column that calculates the height
#       of each layer.
lb <- as.vector(cohort$layer_bottom)
lt <- as.vector(cohort$layer_top)

#2.c) extract out cohort data for 2018.
cohort18 <- cohort %>%
  mutate(depth = lb-lt) %>%
   filter(year == 2018)

cohort19 <- cohort %>%
  mutate(depth = lb-lt) %>%
   filter(year == 2019)

cohort20 <- cohort %>%
  mutate(depth = lb-lt) %>%
   filter(year == 2020)

#2.d) how many layers are exactly one meter
# below the last year of interest (2020) in the depth column?
cohort20 %>%
  summarise(total_depth = sum(depth))
```


```{r}
rm(a,
   elevation,
   lb,
   lt,
   fastomPackingDensity,
   omPackingDensity,
   resp,
   fast,
   slow,
   cohort18,
   cohort19,
   cohort20,
   memCohortExample2)
```



# *Phase 2: PEPRMT*

## *Ecosystem Respiration*

### **Step 1: import your data to environment.**



```{r message=FALSE}
data <- read_csv(file="./Data_sets/EL_2018_2020_master_9_30_20.csv") %>%
  mutate(GI = rep(0, nrow(.)),
         Season = rep(0, nrow(.))) %>%
         #wc90_filler = rep(0, nrow(data))
  select(Day_of_year,
         Discont_day_of_year,
         Air_temp_daily_ave_degree_C,
         Water_table_daily_ave_cm,
         #PAR_daily_ave_umol_m2_day,
         #FPAR,
         GPP_gC_m2_day,
         GI,
         Season,
         #wc90_filler,
         Wetland_age_years,
         Year) %>%
 filter(between(Year, 2018, 2020)) %>%
  as.data.frame(.)
```


### **Step 2: Prepare the data for peprmt Reco script**

In this step, I substitute the tower observed measures of GPP for a vector of zeroes.

Within the PEPRMT Reco function, GPP measures are susbstituted directly fora variable called "C2in", which is then used to calculate the S2 pool of carbon in PEPRMT. By converting the GPP time step measures to zero, the S2 carbon pool depends strictly on the Labiale Carbon Pool, which is the sum of the fast and respired organic matter pools calculated from MEM.

```{r}
# 2.a) substitute the peprmt GPP data with
#       MEM-calculated NPP data; relabel column.
data[5] <-  c(rep(0, nrow(data)))
```


### **Step 3: Execute peprmt Reco using sum of fast and slow OM of MEM for 2018**

```{r}
#Run a loop for each year in a vector of years where you:
#   - extract year data
#   - store in df
#   - run mem-pep Reco on that data
#   - store df as element of list


#3.a) pull your reco function in and setup storage space
source('./Peprmt_files/PEPRMT_final_sys_CO2_Reco_reduced.R')
year_vector <- c(2018,2019,2020)
outcome_lst <- list(rep(NA, length(year_vector)))
daysofyear <- list(c(1, 365), c(366,730), c(731,920))
wetland_type <- "coastal"
     
#3.b) run the loop
  for(i in 1:length(year_vector)) {
    cohort_i <- cohort %>%
      filter(year == year_vector[i]) %>%
      mutate(depth = layer_bottom - layer_top)
    
    # set up your theta values. note that multiply by 10^6 to
    #convert from gC/cm^3 to gC/m^2, when we consider one meter
    #of soil across all years of interest.
    theta5 <- (mean(cohort_i$slow_OM)/mean(cohort_i$depth))*1000000
    theta6 <- ((mean(cohort_i$fast_OM)+mean(cohort_i$respired_OM))/mean(cohort_i$depth))*1000000
   
     # run peprmt Reco
    theta <- c(-1, 0, -1, 0, theta5, theta6)
    ER_mod_RR <- PEPRMT_final_sys_CO2_Reco(data,
                                           theta,                                           wetland_type)
    
    # merge peprmt output + obs data
    co2_i <- ER_mod_RR %>%
      as.data.frame(.) %>%
      mutate(DOY = c(data$Discont_day_of_year),
             DOY_cont = c(data$Day_of_year)) %>%
      select(Reco_full,
             DOY, 
             DOY_cont,#,
             S1,
             S2
             ) %>%
      rename(CO2_flux=Reco_full) %>%
      mutate(data_type = c(rep(paste0("mem_pep_",year_vector[i],sep=""),
                               nrow(.)))) 
    
    #filter the data between min & max values of obs data
    #for given year
    s <- unlist(daysofyear[i])
    co2_i <- co2_i %>%
      slice(s[1]:s[2])
    
    #store your dataframe iteration of Reco into the list
    outcome_lst[[i]] <- co2_i
  }
```


### **Step 4: Munge the data from loop**

```{r message=FALSE}
#4.a) row bind the elements of loop outcomes together.
bigc02 <- do.call('rbind', outcome_lst)

#4.b) save your S1 and S2s into a separate df, then toss.
s1 <- bigc02$S1
s2 <- bigc02$S2

carbon_pools <- as.data.frame(cbind(s1, s2))
colnames(carbon_pools) <- c("S1", "S2")

#4.c) remove S1 and S2 from bigc02.
bigc02 <-bigc02 %>%
  select(-S1,-S2)

#4.d) Bring in the observed Reco data  
data2 <- read_csv(file="./Data_sets/EL_2018_2020_master_9_30_20.csv") %>%
  filter(between(Year, 2018, 2020)) %>%
  select(Reco_gC_m2_day,
         Discont_day_of_year,
         Day_of_year) %>%
  mutate(data_type = c(rep("obs_data", nrow(.)))) %>%
  rename(CO2_flux = Reco_gC_m2_day,
         DOY = Discont_day_of_year,
         DOY_cont= Day_of_year) %>%
  as.data.frame(.)


#4.c) bind loop data with observed data
bigc02 <- bigc02 %>%
  as.data.frame(.) %>%
  #slice(-(272:783)) %>%
  bind_rows(.,data2) %>%
  pivot_wider(names_from = data_type,
              values_from = CO2_flux) %>%
  pivot_longer(c(mem_pep_2018,
                 mem_pep_2019,
                 mem_pep_2020,
                 obs_data),
               names_to = "data_source",
               values_to = "Respiration")
```


### **Step 5: Plot your Reco values**

```{r}
#5.a). return a plot of the model ouputs & obs. flux data
ggplot(data=bigc02,
       mapping = aes(x = DOY_cont,
                     y = Respiration,
                     color = data_source)) +
  geom_line(na.rm=TRUE) +
  labs(x = "Day of year",
       y = expression(paste("Respiration",
                             " ","(grams C"," ",
                             m^{-2},day^{-1},")",)),
       color = "Respiration Source",
       title = "MEM-PEPRMT: Respiration 2018-2020",
       subtitle = "data obtained from Eden Landing") +
  scale_y_continuous(expand=c(0,0)) +
  scale_color_manual(labels = c("Modeled 2018",
                                "Modeled 2019",
                                "Modeled 2020",
                                "Obs"),
                     values=c("red",
                              "blue",
                              "green4",
                              "black")) +
  theme(panel.grid = element_blank(),
        panel.background = element_rect(fill = "white")
        )
```


## *Methane flux*

```{r}
#clear the environment
rm(i,
   s,
   startYear,
   theta,
   theta6,
   theta5,
   wetland_type,
   year_vector,
   PEPRMT_final_sys_CO2_Reco,
   co2_i,
   cohort_i,
   #data2,
   daysofyear,
   s1,
   s2,
   bigc02,
   #data,
   data2
   )
```


### **Step 1: Merge the observed data with the outputs of the 2018-2020 respiration component of PEPRMT**

In order to run the methane component of PEPRMT, the model requires measures of carbon within the labiale and SOC pools, which are identified as columns S1 and S2 of the respiration output.


```{r message=FALSE}
#1.a): bring back your observed data for 2018-2020 with gpps.
# make sure the columns match
data <- read_csv(file="./Data_sets/EL_2018_2020_master_9_30_20.csv")
dt <- read_csv(file="./Data_sets/EL_2018_2020_master_9_30_20.csv") %>%
  mutate(GI = rep(0, nrow(.)),
         Season = rep(0, nrow(.)), 
         wc90_filler = rep(0, nrow(data)) ) %>%
  select(Day_of_year,
         Discont_day_of_year,
         Air_temp_daily_ave_degree_C,
         Water_table_daily_ave_cm,
         PAR_daily_ave_umol_m2_day,
         FPAR,
         GPP_gC_m2_day,
         GI,
         Season,
         wc90_filler,
         Wetland_age_years,
         Year) %>%
 filter(between(Year, 2018, 2020)) %>%
  as.data.frame(.)


#1.b): merge your Reco ouputs with carbon pools and the
#       filtered 2018 data outputed from MEM.
a<-cbind(dt, carbon_pools)
rm(dt)
```


### **Step 2: Load the salinity inputs into the environment**

I will be using a revised version of the peprmt methane component that incorporates salinity measures. The model requires three additional inputs: Sal_int, Sal_slope, and salinity measures.

Note that there is no need for unit conversions applied to the S1 and S2 carbon pools. In the previous conversion in Reco, the pools converted from g C/cm^3 to g C/m^2 when considered with respect to one meter depth. there should be no unit changes between the phases of PEPRMT.

Note: Previously I had been applying a unit conversion to the S1 and S2 pools. I wasn't totally sure why;  I was trying to get the values to approximate a percentage and because without a conversion totally kills the PEPRMT Ch4. The reasons the model crashes is because currently I am applying a random sample of S1 and S2 values around the gamma distribution that is centered around the mean..??

```{r}
#2.a)  generate a salinity vector 
salinity <- c(data$Salinity_daily_ave_ppt)

#2.b) generate the inputa data frame to feed into PEPRMT CH4.
a <- a %>%
  mutate(Sal = salinity) %>%
  select(Day_of_year:Wetland_age_years, Year, Sal, S1, S2)
```


### **Step 3: setup other inputs to Ch4 script**

```{r}
# 3.a) create your model inputs
#theta= c(-10, 3, -10, 2, 10, 3)
theta= c(-10, 2, -10, 2, 10, 2)

Sal_Slope2 <- -0.05 
Sal_int2 <- 0.73585193

Sal_Slope1 <- -0.2 
Sal_int1 <- 8.92208026

#first attempt values below
#Sal_Slope1 <- -0.05 
#Sal_Slope2 <- -0.2 
#Sal_int1 <- -0.2
#Sal_int2 <- -2.2

#original salinity measures 
#Sal_int <- 0.215850539
#Sal_slope <- -0.008947612
```



### **Step 4: Implement the latest version of the methane component of MEM-PEP one time**

Please the see the notes within the file PEPRMT_final_sys_gammaCH4_V5.R for details on model modifications; this is the latest version of the methane component of PEPRMT.

```{r}
#4.a) load the new methane flux script
source('./Peprmt_files/PEPRMT_Final_sys_CH4_V6.R')
```


```{r}
#4.b) execute your model given the inputs
tic()
CH4_mod_RR <- CH4_daily_step(theta,
                             a,
                             Sal_int1,
                             Sal_int2,
                             Sal_Slope1,
                             Sal_Slope2,
                             x = 1,
                             wetland_type = 2)
toc()
beep(1)
```


do a quick plot of the ch4 outputs below

```{r}
test_graph <- CH4_mod_RR
x <- data$CH4_gC_m2_day
#x<- RR_1617_master$daily.wm.gf.obs

test_graph %>%
    mutate(observed = x,
           sum_of_ms = M1 + M2,
           roxi =r_oxi,
           ch4water = ch4_water) %>%
  pivot_longer(c(avg_Pulse_Emission_Total,
                 M1,
                 M2,
                 hydro_flux,
                 Plant_flux,
                 observed,
                 sum_of_ms,
                 roxi,
                 ch4water),
               names_to='outputs',
               values_to='measures'
               ) %>%
  ggplot(data=., mapping = aes(x=index,
                               y=measures,
                               color= outputs)) +
  geom_line()
```


### **Step 5: Munge your data for graphing**

```{r}
#5.a) clear environ
rm(carbon_pools,
   salinity,
   p,
   Sal_int,
   Sal_slope,
   salt,
   theta,
   year_vector,
   CH4_daily_step,
   daysofyear,)
```


```{r message=FALSE}
#5.b) merge your data sets:
#x: matlab observed ch4
#a: matlab data used to extract day counters
#Ch4_mod_RR: peprmt ch4 model output
data_CH4_mod_EL <- cbind(a,
                         CH4_mod_RR,
                         x) %>%
  select(Day_of_year,
         S1,
         S2,
         avg_Pulse_Emission_Total,
         x) %>%
  rename(obs_ch4_matlab = x)
```


Do side step where you calculate your rmse, r squared and cumulative sums of obseved and predicted methane.

```{r}
#5.c 
#cumulative sums
sum(data_CH4_mod_EL$avg_Pulse_Emission_Total)
sum(data_CH4_mod_EL$obs_ch4_matlab)

tail(cumsum(data_CH4_mod_EL$avg_Pulse_Emission_Total))
tail(cumsum(data_CH4_mod_EL$obs_ch4_matlab))

#rmse
o <- data_CH4_mod_EL$obs_ch4_matlab
m <- data_CH4_mod_EL$avg_Pulse_Emission_Total

sqrt(mean((m - o)^2))

#r squared via basic linear model
plot(m~o, col = c("red","blue"))

#r squared
mymodel <- lm(avg_Pulse_Emission_Total~obs_ch4_matlab,data = data_CH4_mod_EL)

summary(mymodel)

plot(mymodel)

```


```{r}
#5.d) one final munge to stack your measures and observed
# measures together.
  data_CH4_mod_EL <- data_CH4_mod_EL %>%
  #  mutate(avg_Pulse_Emission_Total = avg_Pulse_Emission_Total)%>% 
    pivot_longer(c(obs_ch4_matlab,
                   avg_Pulse_Emission_Total#,
                   #obs_ch4_recent
                   ),
                 names_to ="methane_flux",
                 values_to = "CH4_measure")
```  


### **Step 6: Plot model outputs alongside obsreved data**

```{r}  
  #6.a) plot
   ggplot(data = data_CH4_mod_EL) +
     geom_line(aes(x= Day_of_year,
                   y= CH4_measure,
                   color=methane_flux)) +
     labs(x = "Day of year",
          y = expression(paste("CH4 exchange"," ","(grams C"," ", m^{-2},day^{-1},")")),
          color = expression(paste("Pulse"," ","Emission")),
          title = "MEM-PEPRMT: Methane Emissions 2016-2017",
          subtitle = "data obtained from Rush Ranch") +
     scale_y_continuous(expand=c(0,0)) +
     scale_color_manual(labels =c("Modeled",
                                  "Obs"),
                        values=c("blue",
                                 "black")) +
     theme(
       panel.grid = element_blank(),
       panel.background = element_rect(fill = "white")
     )
```


### save certain data for graphing for poster presentations
```{r}
save(CH4_mod_RR,
     a,
     RR_1617_master,
     data,
     salinity,
     file="/home/worker/Documents/AGU_poster_graphs/revised_RushRanch_data.RData")
```
