---
title: "MEM-PEPRMT: analysis with new thetas"
author: "Michael Najarro"
date: "11/10/2020"
output: html_document
---

# *Background*
The MEM-PEPRMT model was created by merging Dr. Jim Morris's  Marsh Equilibrium Model (MEM) - courtesy of Dr. James Holmquists work in developing MEM in R - with Dr. Patty Oikawa's Peatland Ecosystem Photosynthesis Respiration and Methane Transport model (PEPRMT). MEM-PEPRMT was built on data from Rush Ranch Site located in Suisun Bay California.


These two models were originally joined by:

  1. Calculating NPP from the sum of above and below ground biomass, downscaled to a daily measure, was used in place of PEPRMT's GPP and feed into both the Reco and methane components of PEPRMT. 

  2. Slow and fast organic matter pools found within each soil accretion within the required time zone were added as theta values to Reco.
  
  3. later, the NPP was dropped from Reco, replacing the GPP with 0s and using only the sum of organic pools; NPP was still used in Reco.


# *Objective*

In this report I investigate a modification of MEM-PEPRMT, where NDVI is used in place of GPP, rather than calculating NPP from MEM. This change was recommended at the previous CCRCN team meeting. MEM will still be used to obtain the summed pools of carbon. 

In addition, I will apply the modified MEM-PEP model to data collected from Eden Landing, located in Hayward California. This is a new site in which the model was neither built around nor has encountered. The data used in this analysis is centered on 2018 tower data from Eden landing data, which already contains daily NDVI measures for the year.

As of 10/21/20, a current form of the rCTM package exists. Due to run time errors with MEM inputs, I use an older version of MEM for this analysis. I plan to compare the two version's outputs.


```{r}
library(pacman)
p_load(oldrCTM331,
       tidyverse,
       magrittr,
       tinytex,
       R.matlab,
       here,
       dplyr,
       tidyr,
       tibble,
       tinytex,
       knitr,
       tictoc,
       beepr)
```


# *Phase 1: MEM*

### *Step 1: Run MEM with Cohorts*

The code below implements MEM. Note that end year is defaulted to 99 years after the start year and mean sea level initial (MSL0) is set to Mean Sea Level (MSL). 

MEM can take optional inputs regarding core sampling data. core sampling data is available for Rush Ranch at the Coastal Carbon Research Coordination Network's online atlas (https://ccrcn.shinyapps.io/CoastalCarbonAtlas/). However for this analysis, I exclude such data as the data is distributed across multiple sites, distributed across general elevation classifiers. I exclude coremin and coremax values. Note that you need to have 'ompacking density' defined initially to run the 'runmemwithcohorts' function.

Inputs for MEM on Eden Landing can be found at the CCRCN google Drive.

Recent data updates from chamber data can provide a more accurate measure for the Eden Landing Site elevation:

```{r}
#import elevation data to environment.
elevation<- read.table(file="./Data_sets/EDENLAND082620.txt",sep=",")

a <- elevation %>%
  rename(record=V1,
         S1=V2,
         S2=V3,
         elevation=V4,
         edcv=V5) %>%
  slice(-25) %>%
  summarise(elev = mean(elevation)*100) %>%
  pull(.)
```


```{r}
omPackingDensity <- 0.82

memCohortExample2 <- runMemWithCohorts(startYear=2010,
                  rslrT1=0.09, #.9
                  rslrTotal=250,
                  initElv=a,
                  MSL=98.1,
                  MHW=194.7,
                  MHHW=213.7,
                  MHHWS= 287,
                  ssc= .000020,#.000025,#.000030,
                  lunarNodalAmp=2.5,
                  bMax=.90,#0.84,#.90,#.048,
                  
                  # in Holmquist rFullBlownMemDev
                  # zVegMin is negative; exclude - sign.
                  zVegMin=100,
                  zVegMax=350,#250,#original350,
                  zVegPeak=180,                  plantElevationType="orthometric",
                  rootToShoot=2,
                  rootTurnover=0.5,
                  rootDepthMax=30,
                  shape="linear",
                  omDecayRate=0.8,
                  recalcitrantFrac=0.35,#0.2,
                  settlingVelocity=0.275,
                  omPackingDensity=0.82,
                 rootPackingDensity = omPackingDensity,
            mineralPackingDensity=2.43,
                  coreYear = 2050#,
                  #coreDepth = 50
                 # coreMins =coreMins[c(1:26),],
                 # coreMaxs = coreMaxs[c(1:26),]
)


slow <- tapply(memCohortExample2$cohorts$slow_OM,
           memCohortExample2$cohorts$year,
           FUN = mean)
slow
fast <- tapply(memCohortExample2$cohorts$fast_OM,
           memCohortExample2$cohorts$year,
           FUN = mean)
fast
resp <-tapply(memCohortExample2$cohorts$respired_OM,
           memCohortExample2$cohorts$year,
           FUN = mean)
resp
```



## **Step 2: obtain the sum of slow and fast organic matter pools**

At this point I entirely skip the calculation and downscaling of NPP.

Instead I calculate two important values from MEM to input into the respiration component of PEPRMT. These values will directly correspond to the slow and fast organic matter components of each soil cohort from MEM. The sum of these values approximate the carbon stored within the labile and SOC pools described in the PEPRMT model.


The MEM ouputs for carbon -The slow and fast organic measures - will become inputs for the ecosystem respiration component of PEPRMT. Specifically, they will equate to the sum of slow and fast organic matter that is within one meter below the soil surface at and up to the time of interest (includes all years prior, up to, and include year of interest) in accordinance with the PEPRMT model (see Oikawa 2017) to simulate soil sampling of core data. 

Note thatin previous iterations of MEM-PEPRMT in Rush Ranch, we did not consider the 1 meter depth because the total height of the soil cohorts from MEM were within 1 meter.

The relationship between these pools can be described as follows:

```{r}
y<- tribble(
  ~MEM_Variable, ~Equiv._PEPRMNT_C_Pool, ~PEPRMNT_Variable, ~theta_element,
  "Slow_OM", "Soil Organic Carbon (SOC) Pool", "S2", "theta[5]",
  "Fast_OM", "Labile Carbon Pool", "S1", "theta[6]"
)

kable(y)
```


Here I calculate the depth of the soil accreted for 2018 and filter out the 2018 data from MEM outputs. I also check if the soil is within one meter of the surface.

```{r}
# first identify the values of slow and fast om
# that are: 
#   1. within a meter
#   2. within a time interval of the earliest
#       year up to, and including the year of 
#       interest.

#note:
# Within a given year, all rows before the newest
# incremented "age" (n+1th row) are considered past cohorts,
# so only sum the rows within a year of interest!

#2.a) create table cohort up to year 2017.
cohort <- as.data.frame(memCohortExample2$cohorts)
#cores <-- as.data.frame(memCohortExample2$core)

#2.b) create a column that calculates the height
#       of each layer.
lb <- as.vector(cohort$layer_bottom)
lt <- as.vector(cohort$layer_top)

#2.c) extract out cohort data for 2018.
cohort18 <- cohort %>%
  mutate(depth = lb-lt) %>%
   filter(year == 2018)

#2.d) how many layers are exactly one meter
# below the last year of interest (2020) in the depth column?
cohort18 %>%
  summarise(total_depth = sum(depth))
```


# *Phase 2: PEPRMT*

## *Ecosystem Respiration*

### **Step 1: import your data to environment.**

Note that there are two important data sets to load into the environment:

  1. the Eden Landing data for 2018
  
  2. salinity data for eden landing, from 2018-2020 data provided by Patty.

```{r}
#1.a) import the tower data data to environment
EL_2018_master <- readMat(here('./Data_sets/2018_EL_data/EL_2018_master.mat'))

colnames(EL_2018_master$data[[1]]) <- c("daily_DOY_cont",
"daily_DOY_disc",
"daily_TA",
"daily_WT",
"daily_PAR",
"FPAR_1",
"daily_gpp_obs",
"NDVI_placeholder",
"Season_daily_co2_random_error",
"daily_co2_gapfilling_error",
"daily_wetland_age",
"FPAR_2")

EL_obs18 <- EL_2018_master$data[[1]]

EL_obs18 <- EL_obs18 %>%
  as.data.frame(.) %>%
  mutate(daily_DOY_disc = as.integer(daily_DOY_disc),
         daily_DOY_cont = as.integer(daily_DOY_cont))
```
  
  

### **Step 2: Prepare the data for peprmt Reco script**

In this step, I substitute the tower observed measures of GPP for a vector of zeroes.

Within the PEPRMT Reco function, GPP measures are susbstituted directly fora variable called "C2in", which is then used to calculate the S2 pool of carbon in PEPRMT. By converting the GPP time step measures to zero, the S2 carbon pool depends strictly on the Labiale Carbon Pool, which is the sum of the fast and respired organic matter pools calculated from MEM.

```{r}
# 2.a) substitute the peprmt GPP data with
#       MEM-calculated NPP data; relabel column.
EL_obs18[7] <-  c(rep(0, nrow(EL_obs18)))
```


### **Step 3: Execute peprmt Reco using sum of fast and slow OM of MEM for 2018**

Problem: I am not sure what the organic matter pools are measured in. I originally thought that they were measured in gC/cm^3, but I can't find a reason why I thought this. I suspect because of the depth measurement.

According to an email between me, Patty, and Jim regarding units for the organic matter pools, sent on June 24th of 2020 titled *Questions on the Cohort table output from MEM*, I ask Jim the following question:

    1. Iam a bit confused on the age component of the cohorts of soil accretion. My understanding is that the time frame being assessed is a cohort per year, but their is also a column "age" which is incrementing by 1 for each year. is each age essentially a day within a year interval?

    2. My second question is with regard to units. What are the units for the fast and slow organic respiration output columns in the cohorts table?

He replies with:

    1. To answer your first question, the time scale are 1 year intervals. The cohorts table contains records of each cohort over every year of the simulation. Year refers to that slice of time in a simulation. Age refers to the age of the cohort relative to the surface. Every year a new cohort made up of sediment is lain down at the top of the depth profile and each cohort beneath it is aged by 1 year.
    
    2. To answer your second question. All mass units are in grams in both the inputs and outputs right now.


The units for the organic matter pool are not in grams Carbon per cm^3. Further, in the MEM function `addCohort.R`, MEM calculates the depth of each cohort using fast adn slow Organic Matter variables. 

The units are for sure include grams Carbon, but possibly may include per year. Morris et al. 1986 has organic matter pools in $\frac{g C}{m^{2}}$. The wikipedia page for Dissolved organic carbon describes various carobn pools in units of  $Tg Carbon$, as well as $\frac{Tg C}{year}$.

Given that MEM is looking at cohorts within each year, and that in PEPRMT the units for C1_init and C2_init -which are esssentially the theta values that come from the MEM organic matter- are in $\frac{gram C}{m-3}$ which converts to  $\frac{gram C}{m-2}$, The units for the organic matter pools must be in:
$$\frac{grams \space Carbon}{year}$$

To convert the grams Carbon per year into grams Carbon per meter squared, I must divide the grams Carbon by the depth measure in cubic centimeters, and then multiply by $\frac{100^3 cm^{3}}{m^3 year}$, which will convert to $\frac{100^3 cm^{3}}{m^2 year}$ when observing data that falls within one meter from the surface for a given year of interest.


~original rational below~:
I multiply the sum of the decayed organic matterpools by 10^6 because the units for each pool's decay rate is at gC/cm^3 in MEM, while in PEPRMT, the units for theta are in gC/m^2. Multiplying gC/cm^3 by 100^3cm^3 will produce gC/m^2. We can bridge the gap between these two units when considering, as stated in Oikawa et al. 2017, that PEPRMT looks at only the first meter below the marsh soil surface. Thus the total depth of the soil needs to be considered. Note that MEM outputs its units typically in grams and centimeters.

Note that the respiration component of peprmt is producing heterotrophic respiration, not total ecosystem respiration (both hetero and autotroph).

```{r}
#2.a) sum the slow_om and fast_om values for theta
# note: with original settings, theta5 and 6 multiplied by 78
# sets theta5 to 57181.228 and theta6 to 5795.37515
theta5 <- (mean(cohort18$slow_OM)/mean(cohort18$depth))*1*10^6
theta6 <- ((mean(cohort18$fast_OM)+mean(cohort18$respired_OM))/mean(cohort18$depth))*(1*10^6)


#theta5<-theta5*78
#theta6<-theta6*78
#theta5 <- sum(cohort18$slow_OM)*(1*10^6)
#theta6 <- (sum(cohort18$fast_OM)+sum(cohort18$respired_OM))*(1*10^6)

theta <- c(-1, 0, -1, 0, theta5, theta6)

#2.b) run reco
source('./Peprmt_files/PEPRMT_final_sys_CO2_Reco.R')
PEPRMT_final_sys_CO2_Reco(EL_obs18,
                          theta,
                          wetland_type="coastal")
```


### **Step 3: Plot the peprmt Reco ouputs relative to observed data**

```{r message=FALSE}
#3.a) load reco model results
load(file="./Data_sets/CH4_input.Rdata")

#3.b) pull out obsreved Reco from matlab data and recent EL data #     from Patty + day counter data.
rd <- read_csv(file="./Data_sets/EL_2018_2020_master_9_30_20.csv")

d <- rd %>%
  filter(Year ==2018)%>%
  select(Reco_gC_m2_day) %>%
  pull(.) %>%
  c(0,.)

b<- c(EL_2018_master$daily.er.obs)
days <- c(EL_obs18$daily_DOY_cont)

#3.c) munge the model ouput data to include observed data +
#     day counters.
Reco_output <- Reco_output %>%
  mutate(obs_Reco_matlab = b,
         #obs_Reco_recent = d,
         Day_of_year=days) %>%
  rename(MEM_PEPRMT_Reco = Reco_full) %>%
  pivot_longer(c(MEM_PEPRMT_Reco,
                 obs_Reco_matlab,
                 #obs_Reco_recent
                 ),
               names_to= "Respiration",
               values_to ="Measures")

#3.d) now plot data
ggplot(data = Reco_output) +
  geom_line(mapping=aes(x=Day_of_year, y= Measures, color=Respiration)) +
   labs(x = "Day of year",
        y = expression(paste("Respiration",
                             " ","(grams C"," ",
                             m^{-2},day^{-1},")",)),
                             #" ",g^{3},m^{2},day^{-1})),
        color = "Respiration Source",
        title = "MEM-PEPRMT: Respiration 2018",
        subtitle = "data obtained from Eden Landing") +
   #scale_fill_discrete(name ="Source", labels = "MEM-PEPRMT", "Obs") +
   scale_y_continuous(expand=c(0,0)) +
   scale_color_manual(labels =c("Modeled", "Obs"),
                      values=c("blue",
                               "black"#,
                               #"red"
                               )) +
   theme(
     panel.grid = element_blank(),
     panel.background = element_rect(fill = "white")
   )
```

At this step it is important to evaluate the Reco results so far. Within the data file "ELM-1812_wc-porosity_updated.xlsx the estimated percent of organic matter found within the Eden Landing sample core data is 15.25% per core. Cutting this pecentage in half would indicate a Carbon ratio of 7.63%. 

Applying this percntage to the bulk density (converted to g/m^3, which is 820821.88) of the cores indicates 125175.34 of the bulk density is organic matter. Assuming that approximately half of the organic matter percentage is comprised of carbon leaves us with 62,587.67 grams of Carbon/m^3.

In theory the Reco output of PEPRMT should approximate the Carbon quantity calculated previously.   


## *Methane flux*

```{r}
#clear the environment
rm(lb,
   lt,
   omPackingDensity,
   theta5,
   theta6,
   theta,
   PEPRMT_final_sys_CO2_Reco,
   elevation,
   cohort,
   #theta,
   #memCohortExample2,
   a,
   b,
   d,
   days
   )
```


### **Step 1: Merge the observed data with the outputs of the 2016 respiration component of peprmt**

In order to run the methane component of PEPRMT, the model requires measures of carbon within the labiale and SOC pools, which are identified as columns S1 and S2 of the respiration output.

In previous runs of MEM-PEPERMT, the model ran the methane component using NPP measures from MEM, downscaled to a daily timestep measure in place of GPP.

Here I now test whether NDVI works as a proxy for GPP, by substituting NDVI in place of GPP.

```{r}
#1.a): reload data from Reco model
load(file="./Data_sets/CH4_input.Rdata")

#1.b): merge your Reco ouputs with the filtered 2018 data
#         outputed from MEM.
a<-cbind(EL_obs18, Reco_output)
p<- c(EL_2018_master$daily.gpp.obs)

a <- a %>%
  select(daily_DOY_cont,
         daily_DOY_disc,
         daily_TA,
         daily_WT,
         daily_PAR,
         FPAR_1,
         daily_gpp_obs,
        NDVI_placeholder,
        #NDVI_placeholder,
         #daily_gpp_obs,
         Season_daily_co2_random_error,
         daily_co2_gapfilling_error,
         daily_wetland_age,
         FPAR_2,
         S1,
         S2) %>%
  #rename(ndvi_inplace_gpp=NDVI_placeholder) %>%
  mutate(daily_gpp_obs = p)
```


### **Step 2: Load the salinity inputs into the environment**

I will be using a revised version of the peprmt methane component that incorporates salinity measures. The model requires three additional inputs: Sal_int, Sal_slope, and salinity measures.

Note that there is no need for unit conversions applied to the S1 and S2 carbon pools. In the previous conversion in Reco, the pools converted from g C/cm^3 to g C/m^2 when considered with respect to one meter depth. there should be no unit changes between the phases of PEPRMT.

Note: Previously I had been applying a unit conversion to the S1 and S2 pools. I wasn't totally sure why;  I was trying to get the values to approximate a percentage and because without a conversion totally kills the PEPRMT Ch4. The reasons the model crashes is because currently I am applying a random sample of S1 and S2 values around the gamma distribution that is centered around the mean..??

```{r}
#2.a) load in to the 2018-2020 EL data to obtain salinity for'18
salinity <- read.csv(file=paste("./Data_sets/EL_2018_2020_master_9_30_20.csv",sep=""))

#2.b) generate a salinity vector.
salt <- salinity %>%
  filter(Year == 2018) %>%
  select(Salinity_daily_ave_ppt) %>%
  pull() %>%
  c(38,.)

#2.c) generate the inputa data frame to feed into PEPRMT CH4.
a <- a %>%
  mutate(Sal = salt) %>%
  select(daily_DOY_cont:FPAR_2, Sal, S1, S2) %>%
  mutate(S1 = S1,#*.0000001,
         S2 = S2# *.0000001
         )
```


### **Step 3: setup other inputs to Ch4 script**

```{r}
# 3.a) create your model inputs
theta= c(-9, 2, -9, 2, 6, 2)
#theta= c(-4, 2, -4, 2, 6, 2)
#theta= c(-5, 1, -5, 1, 5, 1)
Sal_int <- 0.215850539
Sal_slope <- -0.008947612
```


### **Step 4: Implement the latest version of the methane component of MEM-PEP one time**

Please the see the notes within the file PEPRMT_final_sys_gammaCH4_V4.R for details on model modifications; this is the latest version of the methane component of PEPRMT.

```{r}
#4.a) load the new methane flux script
source('./Peprmt_files/PEPRMT_Final_sys_CH4_V5.R')
```

```{r}
#21.b) execute your model given the inputs
tic()
CH4_mod_RR <- CH4_daily_step(theta,
                             a,
                             Sal_slope,
                             Sal_int,
                             x = 1,
                             wetland_type = 2)
toc()
beep(1)
```


do a quick plot of the ch4 outputs below

```{r}
test_graph <- CH4_mod_RR
x<- EL_2018_master$daily.ch4.obs

test_graph %>%
    mutate(observed = x,
           sum_of_ms = M1 + M2,
           roxi =r_oxi,
           ch4water = ch4_water) %>%
  pivot_longer(c(avg_Pulse_Emission_Total,
                 M1,
                 M2,
                 hydro_flux,
                 Plant_flux,#,
                 observed,
                 sum_of_ms,
                 roxi,
                 ch4water),
               names_to='outputs',
               values_to='measures'
               ) %>%
  ggplot(data=., mapping = aes(x=index, y=measures, color= outputs)) +
  geom_line()
```


### **Step 5: Plot model outputs alongside obsreved data**

```{r}
#5.a) clear environ
# rm(#Reco_output,
#    salinity,
#    p,
#    Sal_int,
#    Sal_slope,
#    theta,
#    salt)
```



```{r message=FALSE}
#5.b) merge your data sets:
#w: matlab observed ch4
#a: matlab data used to extract day counters
#Ch4_mod_RR: peprmt ch4 model output
# rd: recent Eden landing tower data from 2018-2020;contains
#     methane tower data.

rd <- read_csv(file="./Data_sets/EL_2018_2020_master_9_30_20.csv")
d <- rd %>%
  filter(Year ==2018)%>%
  select(CH4_gC_m2_day) %>%
  pull(.) %>%
  c(0,.)

w <- c(EL_2018_master$daily.ch4.obs)

data_CH4_mod_EL <- cbind(a,CH4_mod_RR,w,d) %>%
  select(daily_DOY_cont,
         S1,
         S2,
         avg_Pulse_Emission_Total,
         w,
         #d
         ) %>%
  rename( obs_ch4_matlab = w#,
          #obs_ch4_recent = d
          )

#data_CH4_mod_EL <- data_CH4_mod_EL[,c(1,14,15,17,18)]
#data_CH4_mod_EL<-rename(data_CH4_mod_EL, obs_ch4 = w)

#5.c) unit change on the microsite column
# multiplied by 1000 to get from micro to nano moles.
  data_CH4_mod_EL <- data_CH4_mod_EL %>%
    mutate(avg_Pulse_Emission_Total = avg_Pulse_Emission_Total)%>% #*-1000) %>%
    pivot_longer(c(obs_ch4_matlab,
                   avg_Pulse_Emission_Total#,
                   #obs_ch4_recent
                   ),
                 names_to ="methane_flux",
                 values_to = "CH4_measure")
  
  #5.d) plot
   ggplot(data = data_CH4_mod_EL) +
     geom_line(aes(x= daily_DOY_cont,
                   y= CH4_measure,
                   color=methane_flux)) +
     labs(x = "Day of year",
          y = expression(paste("CH4 exchange"," ","(grams C"," ", m^{-2},day^{-1},")")),
          color = expression(paste("Pulse"," ","Emission")),
          title = "MEM-PEPRMT: Methane Emissions 2018",
          subtitle = "data obtained from Eden Landing") +
     scale_y_continuous(expand=c(0,0)) +
     scale_color_manual(labels =c("Modeled",
                                  "Obs"),
                        values=c("blue",
                                 "black"#,"grey"
                                 )) +
     theme(
       panel.grid = element_blank(),
       panel.background = element_rect(fill = "white")
     )
```


### save data sets for poster graphs

```{r}
save(CH4_mod_RR,
     salinity,
     a,
     file="/home/worker/Documents/AGU_poster_graphs/revised_EL_data.RData")
```

